---
title: "Coursera statinference-006 course project"
author: "Raymond Nelson"
date: "September 12, 2014"
output: html_document
---

# Simulation of an Exponential Distribution

## Assignment

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. 

The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda. 

Set lambda = 0.2 for all of the simulations. 

In this simulation, you will investigate the distribution of averages of 40 exponential(0.2)s. Note that you will need to do a thousand or so simulated averages of 40 exponentials.

## Instructions

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponential(0.2)s.  

You should

1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.
2. Show how variable it is and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.
4. Evaluate the coverage of the confidence interval for 1/lambda: .

## Preliminaries


First set the working directory.

```{r}
setwd("~/Documents/RAYMOND/Coursera/Statistical Inference 9-1-2014 to 9-29-2014/statInfProject")
```

Then construct the replications of the simulated distribution of 1000 iterations of 40 random exponential variables with lambda = .2

```{r}
numVars <- 40
numSim <- 1000
lambda <- .2

# construct the NULL matrix to hold the data
# 40 columns represent random exp values
# 1000 rows will hold the results of the simulation run

set.seed(1234)
simMatrix <- matrix(rexp(numSim * numVars, lambda), numSim)
```

## Project questions


### 1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.

```{r}
# using rowMeans
rm <- rowMeans(simMatrix)
m <- mean(rm)

# another method using apply
m <- mean(apply(simMatrix, 1, mean))

# the theoretical center of the distribution is 1 / lambda
tm <- 1 / lambda
```

### 2. Show how variable it is and compare it to the theoretical variance of the distribution.

```{r}
# calculation of the sampling SD is done with apply 
s <- mean(apply(simMatrix, 1, sd))
# standard deviation
s
# variance
s^2

# the theoretical SD of the distribution is 1 / lambda for which the square is the variance.
tsd <- 1 / lambda
tsd^2
```

### 3. Show that the distribution is approximately normal.

```{r}
# histogram to evaluate normality of the distribution
hist(rowMeans(simMatrix), breaks = 30)
```

Another example with ggplot.

```{r}
library(ggplot2)
rm <- data.frame(rowMeans(simMatrix))
ggplot(rm, aes(x=rowMeans.simMatrix.)) + 
        geom_histogram(aes(y=..density..),
                       binwidth=.5,
                       colour="black", fill="white") +
        geom_density(alpha=.2, fill="pink")           
```

### 4. Evaluate the coverage of the confidence interval for 1/lambda: .

```{r}
# first calculate the standard error of the SD of the simulated distribution
# upper limit
ul <- s + 1.9599 * (s / sqrt(numSim))
ul        
# lower limit        
ll <- s - 1.9599 * (s / sqrt(numSim))
ll
```


